---
title: "Marketing Spend vs Purchase Behavior Analysis"
author: "Nattamon Wattanapenpaiboon"
format: 
    html:
        theme: cerulean
    pdf: default
editor: visual      
---

```{r}
#| include: false
# loading libraries
library(tidyverse)
library(knitr)
library(lubridate)
library(ggplot2)
# load raw data
raw_data <- read_csv("data/36798800_Nattamon Wattanapenpaiboon.csv")
```

```{r}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

In the increasingly competitive online retail sector, the efficient allocation of marketing budgets is paramount to sustainable growth. During the last fiscal quarter, our team executed a personalized campaign where expenditures on advertisements, emails, and coupons were varied at the individual customer level. This report analyzes the resulting performance data to evaluate the direct impact of these targeted investments on our bottom line. By establishing the link between expenditure and revenue, we can more effectively refine our customer segmentation strategies for future promotional cycles.

# Research Question

The primary objective of this investigation is to determine whether an increase in marketing spend per customer results in a proportional rise in their total purchase amounts. Specifically, we will examine if our high-spend segments consistently outperform lower-investment groups in terms of total sales volume. This analysis will confirm whether our current personalized spending model successfully incentivizes higher consumer expenditure.

## Dataset Introduction

```{r}
# Data cleaning
# Clean customer_id
marketing_data <- raw_data |> mutate(customer_id = str_remove(customer_id, "^.{9}"),
# Clean and convert signup_date string to date format
signup_date = parse_date_time(signup_date, orders = c("dmy", "mdy", "ymd")),
                                         signup_date = as.Date(signup_date))
```

The marketing campaign dataset used in this analysis serves as the foundational record of our personalized outreach strategies across ads, emails, and coupons during the previous quarter. It tracks key individual-level metrics, including unique customer_id, their acquisition timeline (signup_date), and our specific financial investment (marketing_spend). We also measure behavioral engagement through visits and total purchase_amounts to quantify the financial return on each marketing dollar. Furthermore, the inclusion of segment and campaign notes provides critical context for distinguishing which demographics or outreach methods were most effective. Collectively, these variables allow for a granular analysis of return on investment and the optimization of future budget allocations.

It consists of \`r ncol(marketing_data)\` variables as shown in the table below, and \`r nrow(marketing_data)\` observations.

```{r}
data_dict <- tibble(
    Variable = names(marketing_data),
    Type = sapply(marketing_data, class)) |>
        mutate(Description = case_when(
            Variable == "customer_id" ~ "Unique customer identification",
            Variable == "signup_date" ~ "Date of customer registration",
            Variable == "marketing_spend" ~ "Total campaign expenditure per customer ($)",
            Variable == "visits" ~ "Count of customer site visits during the campaign",
            Variable == "purchase_amount" ~ "Total revenue generated from the customer ($)",
            Variable == "segment" ~ "Customer segment i.e. Enterprise, SMB, Consumer",
            Variable == "notes" ~ "Qualitative record of campaign outreach outcomes"
        ))

kable(data_dict)
```

Create two other summaries of your data, one as a plot and one as a table. Examples include counts of certain variables in groups, or summary statistics of numeric variables. Make sure these summaries are providing a broader context to the report and are not just arbitrary commentaries.

# Results

- Create 2 figures maximum of your data that will help you answer your research question. Each figure must have a caption using the options inside the R code chunk. Create the figures using the `ggplot2` package.

- The figures must have captions, and must be referred to in the text of the results section.

As well as the figures, you must also include a paragraph describing these results and how they link back to your research question. Remember, you can use inline code to extract certain numbers automatically.