---
title: "Marketing Spend vs Purchase Behavior Analysis"
author: "Nattamon Wattanapenpaiboon"
format: 
    html:
        theme: cerulean
    pdf: default
editor: visual      
---

```{r}
#| include: false
# loading libraries
library(tidyverse)
library(knitr)
library(lubridate)
library(ggplot2)
# load raw data
raw_data <- read_csv("data/36798800_Nattamon Wattanapenpaiboon.csv")
```

```{r}
#| include: false
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Introduction

In the increasingly competitive online retail sector, the efficient allocation of marketing budgets is paramount to sustainable growth. During the last fiscal quarter, the company executed a personalized campaign where expenditures on advertisements, emails, and coupons were varied at the individual customer level. This report analyzes the resulting performance data to evaluate the direct impact of these targeted investments on our income. By establishing the link between expenditure and revenue, we can more effectively refine our customer segmentation strategies for future promotional cycles.

# Research Question

The primary objective of this investigation is to determine if higher marketing expenditure per customer directly results in increased purchase amounts. Additionally, we will analyze whether the responsiveness to this personalized spend varies significantly across our customer segments. By establishing this link, we can evaluate the efficiency of our current model and identify which customer groups yield the highest return on investment.

# Dataset Introduction

The dataset provides a comprehensive record of our personalized outreach strategies. It tracks individual customer acquisition timelines and our specific financial investment alongside behavioral metrics such as site visits, campaign outreach notes, and total revenue generated. The inclusion of customer segments provides the necessary context to evaluate how responsiveness to marketing investment varies across different business tiers. Collectively, these variables allow us to quantify the financial return on each marketing dollar spent and identify opportunities for budget optimization.

**Data Cleaning**\
To make the data ready for analysis, I first cleaned the customer IDs by removing the extra prefix digits from each entry. I also unified the sign-up dates into a single, consistent format so they could be easily tracked over time. I also replaced missing values with \$0.00 for purchase amount to show that the customers have not made any purchases. Lastly, I updated the campaign notes to ensure that any "N/A" entries were correctly recognized as missing information.

```{r}
# Data cleaning
# Remove student id from customer_id
marketing_data <- raw_data |> mutate(
  customer_id = str_remove(customer_id, "^.{9}"),
# Convert signup_date string to date format
  signup_date = parse_date_time(signup_date, orders = c("dmy", "mdy", "ymd")),
  signup_date = as.Date(signup_date),
# Remove "N/A" from notes values and replace missing values with 0 for purchase_amount
  notes = na_if(notes, "N/A"), 
purchase_amount = replace_na(purchase_amount, 0.00))
```

## Dataset Description

The marketing campaign dataset consists of `r ncol(marketing_data)` variables and `r nrow(marketing_data)` observations. The variables capture customer identity, engagement, and financial performance. Customer identifiers and categorical fields, such as business segment and campaign notes, are stored as character variables. The sign-up date is recorded as a Date object to reflect the timing of customer acquisition. The dataset also includes quantitative metrics, with numeric variables for marketing spend and purchase amount, along with a count of site visits to measure individual engagement.

```{r}
#| label: fig-customer_tenure
#| fig-cap: Customer Aquisition Timeline
marketing_data |> 
    filter(!is.na(signup_date)) |> 
    ggplot(aes(x = signup_date)) +
    geom_histogram(bins = 20, fill = "#3498db", colour = "white") +
    theme_minimal() +
    labs(x = "Sign-up Date", y = "Number of New Customers")
```

@fig-customer_tenure illustrates the acquisition timeline of our customers, providing essential context on the maturity of our target audience. This distribution shows that our campaign reached both long-standing loyalists and recently acquired users, which is important because a customer's tenure often influences how they respond to personalized marketing spend.

```{r}
#| label: tbl-summary_by_segment
#| tbl-cap: "Comparative Summary of Marketing Investment and Customer Engagement by Segment"
marketing_data |>
group_by(segment) |>
summarise(
    "Customer Count" = n(),
    "Avg. Marketing Spend (\\$)" = mean(marketing_spend),
    "Avg. Purchase (\\$)" = mean(purchase_amount),
    "Avg. Visits" = mean(visits)) |>
    mutate(across(where(is.numeric), ~ round(.x ,2)))|>
    kable()
```

@tbl-summary_by_segment outlines the strategic allocation of our campaign resources across three distinct business channels. The Consumer segment constitutes the majority of our targeted base and received the highest average financial investment per customer. However, the Enterprise tier, despite being our smallest group, demonstrated the highest levels of brand engagement through site visits and generated the strongest average revenue conversion. Meanwhile, the SMB segment represents a more conservative investment tier, resulting in the lowest relative revenue returns compared to the other groups. This variation in how different business tiers respond to our marketing expenditure provides essential baseline context for evaluating the campaign's overall performance. Nevertheless, these average figures only tell a partial story, as they are susceptible to being skewed by individual high-value outliers. Consequently, while Table 2 offers a helpful overview of the dataset's scale and diversity, a deeper analysis of the distribution and individual customer responsiveness is required in the following results to reach a more robust conclusion regarding the campaign's true efficiency.

# Results

Analysis of @fig-correlation reveals a clear positive relationship between the company's marketing expenditure and customer purchase behavior with a calculated correlation of `r round(cor(marketing_data$marketing_spend, marketing_data$purchase_amount, use = "complete.obs"), 2)`. This confirms that our strategy of increasing spend successfully drives higher sales volumes across the board. Notably, because our customer base includes a wide range of signup dates, these results confirm that our personalized marketing strategy is effective for both long-term loyalists and newly acquired customers alike. However, @fig-efficiency reveals a critical strategic challenge: despite this positive correlation, the median efficiency across all segments remains below the 1.0 break-even threshold (shown by the red dashed line). This indicates that while we are successfully "moving the needle" on revenue, the cost of acquisition for the typical customer currently outweighs the immediate financial return. The data further reveals that this "spend-to-purchase" relationship is highly segment-dependent. The Enterprise segment is the most predictable but least reactive, with a tight grouping of data points sitting furthest below the profitability line, suggesting that simply increasing spend in this tier may not be financially worth the investment under the current model. In contrast, the SMB and Consumer segment display greater variability and remain closer to the break-even median, representing our most viable path toward a self-sustaining return on investment ratio. Ultimately, Consumer segment is uniquely characterized by extreme high-value outliers; while the "average" consumer remains below the break-even point, these superstar performers suggest that our personalized spend has the highest potential for explosive, multi-fold returns within the retail space. Collectively, these findings suggest that while our current campaign is effective at generating higher purchase amounts, it requires significant optimization for profitability.

```{r}
#| label: fig-correlation
#| fig-cap: Relationship Between Marketing Spend and Purchase Amount by Segment
#| fig-pos: "H"

ggplot(marketing_data, aes(x = marketing_spend, y = purchase_amount, colour = segment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se= FALSE, colour = "black") +
  theme_minimal() +
  labs(
    x = "Marketing Spend ($)",
    y = "Purchase Amount ($)"
  )
```

```{r}
#| label: fig-efficiency
#| fig-cap: Analysis of Purchase-to-Spend Efficiency Across Marketing Segments
#| fig-height: 6
#| fig-width: 6
#| fig-pos: "H"

purchase_spend_ratio <- marketing_data$purchase_amount / marketing_data$marketing_spend

ggplot(marketing_data, aes(x = segment, y = purchase_spend_ratio, fill = segment)) +
  geom_boxplot() +
  geom_hline(yintercept = 1.0, linetype = "dashed", colour = "red") +
  theme_minimal() +
  labs(
    x = "Customer Segment",
    y = "Purchase to Spend Ratio"
  )
```
