---
title: "Marketing Spend vs. Purchase Behavior Analysis"
author: Nattamon Wattanapenpaiboon
format: 
    html:
        theme: cerulean
    pdf: default
 editor: visual      
---

```{r}
#| include: false
# loading libraries
library(tidyverse)
library(knitr)
library(lubridate)
library(ggplot2)
# load raw data
raw_data <- read_csv("data/36798800_Nattamon Wattanapenpaiboon.csv")
```

# Introduction

In the increasingly competitive online retail sector, the efficient allocation of marketing budgets is paramount to sustainable growth. During the last fiscal quarter, our team executed a personalized campaign where expenditures on advertisements, emails, and coupons were varied at the individual customer level. This report analyzes the resulting performance data to evaluate the direct impact of these targeted investments on our bottom line. By establishing the link between expenditure and revenue, we can more effectively refine our customer segmentation strategies for future promotional cycles.

# Research Question

The primary objective of this investigation is to determine whether an increase in marketing spend per customer results in a proportional rise in their total purchase amounts. Specifically, we will examine if our high-spend segments consistently outperform lower-investment groups in terms of total sales volume. This analysis will confirm whether our current personalized spending model successfully incentivizes higher consumer expenditure.

## Dataset Introduction

```{r}
# Clean customer_id
marketing_data <- raw_data |> mutate(customer_id = str_remove(customer_id, "^.{9}"),
# Clean and convert signup_date string to date format
signup_date = parse_date_time(signup_date, orders = c("dmy", "mdy", "ymd")),
                                         signup_date = as.Date(signup_date))

```

size of the dataset - number of observations, variables, variable types

- The dataset shows a record of a personalized marketing campaign.
- It tracks each customer with their sign up history, the company's marketing investment, and customer's engagement, and the final revenue. It also includes a segment variable to categorize customers.

The marketing campaign dataset used in this analysis consists of \`r nrow(marketing_data)\` observations and \`r ncol(marketing_data)\` variables.

Create two other summaries of your data, one as a plot and one as a table. Examples include counts of certain variables in groups, or summary statistics of numeric variables. Make sure these summaries are providing a broader context to the report and are not just arbitrary commentaries

# Results

- Create 2 figures maximum of your data that will help you answer your research question. Each figure must have a caption using the options inside the R code chunk. Create the figures using the `ggplot2` package.

- The figures must have captions, and must be referred to in the text of the results section.

As well as the figures, you must also include a paragraph describing these results and how they link back to your research question. Remember, you can use inline code to extract certain numbers automatically.